---
title: "Notebook preparing semi-synthetic data for simulation"
author:
  - Bénédicte Colnet [Inria, Paris-Saclay]
date: "June 2022"
output:
  html_document:
    code_folding: "hide"
    number_sections: no
    toc: yes
    toc_depth: 2
  pdf_document:
    toc: yes
abstract: | 
  This script is used only once to prepare the semi-synthetic simulation where the covariates mimic a real-life case for traumatic brain injured patients using data from CRASH-3 and the Traumabase. 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = FALSE, verbose = FALSE, message = FALSE, warning = FALSE)

# Set random generator seed for reproducible results
set.seed(123)

# Libraries
library(readxl) # Read xlsx
library(dplyr)
library(ggplot2)
library(scales) # for the bar plot with the percentage


# covariates set considered
RELEVANT_COVARIATES = c("time_to_treatment", "gender", "age", "pupilReact_num",  "Glasgow.initial", "systolicBloodPressure")
```

# CRASH3

## Load CRASH-3 data (RCT)

```{r, message=F, results='hide'}
# Load CRASH3 data
rawData_CRASH3 <- read_xlsx("~/Documents/data/crash/crash-data/crash3/CRASH-3_dataset_anonymised_for_Freebird.xlsx", na =c("", "NR", "NA", "NF","IMP", "ND"),)
```


Trial data contains missing values, and in particular some rows with completely missing values. We remove them.

```{r}
rawData_CRASH3[rowSums(is.na(rawData_CRASH3[ , 1:6])) == 6, ]
```


```{r}
rawData_CRASH3 <- rawData_CRASH3[rowSums(is.na(rawData_CRASH3[ , 1:6])) != 6, ]
rawData_CRASH3 <- rawData_CRASH3[rowSums(is.na(rawData_CRASH3[ , 1:6])) != 5, ]
```


## Covariates preparation or recoding or renaming

```{r}
# GSC
rawData_CRASH3$Glasgow.initial <- as.numeric(substring(rawData_CRASH3$gcsEyeOpening, 1, 1)) + 
                                 as.numeric(substring(rawData_CRASH3$gcsMotorResponse, 1, 1)) +
                                 as.numeric(substring(rawData_CRASH3$gcsVerbalResponse, 1, 1))
 
# Women (1) and men (0)
rawData_CRASH3$gender <- rawData_CRASH3$sex
rawData_CRASH3$gender <- ifelse(rawData_CRASH3$gender == "Female", 1, 0)

# Pupil reactivity
rawData_CRASH3$pupilReact_num <- case_when(rawData_CRASH3$pupilReact == "Both React" ~ 2,
                                           rawData_CRASH3$pupilReact == "One Reacts" ~ 1,
                                           rawData_CRASH3$pupilReact == "None React" ~ 0,
                                           rawData_CRASH3$pupilReact == "Unable to assess" ~ -1) 

# Time between injury and treatment
## data treatement from string to numeric hours and minutes
rawData_CRASH3$timeSinceInjury_h = format(as.POSIXct(rawData_CRASH3$timeSinceInjury, format="%Y-%m-%d %H:%M"), format="%H")
rawData_CRASH3$timeSinceInjury_h <- as.numeric(rawData_CRASH3$timeSinceInjury_h)
rawData_CRASH3$timeSinceInjury_m = format(as.POSIXct(rawData_CRASH3$timeSinceInjury, format="%Y-%m-%d %H:%M"), format="%M")
rawData_CRASH3$timeSinceInjury_m <- as.numeric(rawData_CRASH3$timeSinceInjury_m)
rawData_CRASH3$time_to_treatment <- rawData_CRASH3$timeSinceInjury_h*60 + rawData_CRASH3$timeSinceInjury_m
```


## Keeping only relevant covariates 

```{r}
CRASH3 <- rawData_CRASH3[, RELEVANT_COVARIATES]
```


## Trial data imputation

CRASH-3 contains missing values.

```{r}
summary(CRASH3)
```


We impute them with the package `missRanger`.

```{r}
library(missRanger)
CRASH3.imputed <- missRanger(CRASH3, num.trees = 100)
```

```{r}
CRASH3.imputed$gender <- as.factor(CRASH3.imputed$gender)
CRASH3.imputed$pupilReact_num <- floor(CRASH3.imputed$pupilReact_num)
CRASH3.imputed$pupilReact_num <- as.factor(CRASH3.imputed$pupilReact_num)
CRASH3.imputed$Glasgow.initial <- floor(CRASH3.imputed$Glasgow.initial )
CRASH3.imputed$Glasgow.initial <- as.factor(CRASH3.imputed$Glasgow.initial)
```


```{r}
summary(CRASH3.imputed)
```



# Traumabase

## Load Target population data data (Observational)

```{r}
# load raw data 
rawData_Traumabase <- read.csv("~/Documents/data/traumabase/data/4_computed_dataset.csv")
print(paste0("Raw traumabase data contains following number of observations: ", nrow(rawData_Traumabase)))
```


## Covariates

```{r}
# Glasgow
rawData_Traumabase$Glasgow.initial <- ifelse(rawData_Traumabase$Glasgow.initial == "ND" | is.na(rawData_Traumabase$Glasgow.initial), NA, as.integer(rawData_Traumabase$Glasgow.initial))

# Age
rawData_Traumabase$age <- rawData_Traumabase$Age.du.patient..ans

# Note that there are two outliers with age>120 years. By manual inspection, we can correct these observations
rawData_Traumabase$age[which(rawData_Traumabase$age==721)] <- 72
rawData_Traumabase$age[which(rawData_Traumabase$age==184)] <- 18

nrow(rawData_Traumabase)
# Keep only TBI patient
rawData_Traumabase$TBI <- ifelse(rawData_Traumabase$Trauma.crânien..lésion.cérébrale.TDM. == "Oui" | rawData_Traumabase$ISS....Head_neck >= 2, 1, 0)
rawData_Traumabase <- rawData_Traumabase[rawData_Traumabase$TBI == 1, ]
nrow(rawData_Traumabase)


# Systolic blood pressure
rawData_Traumabase$systolicBloodPressure <- rawData_Traumabase$Pression.Artérielle.Systolique..PAS..à.l.arrivée.du.SMUR 

# Women (1) and men (0)
rawData_Traumabase$gender <- ifelse(rawData_Traumabase$Sexe == "Féminin", 1, 0)

# Pupil reactivity
x <- rawData_Traumabase[,"Anomalie.pupillaire..Pré.hospitalier."]

rawData_Traumabase$pupilReact <- case_when(x == "Non" ~ "Both React",
                                           x == "Anisocorie (unilatérale)" ~ "One Reacts",
                                           x == "Mydriase Bilatérale" ~ "None React",
                                           x == "Pas précisé" ~ "Unable to assess") 

rawData_Traumabase$pupilReact_num <- case_when(rawData_Traumabase$pupilReact == "Both React" ~ 2,
                                           rawData_Traumabase$pupilReact == "One Reacts" ~ 1,
                                           rawData_Traumabase$pupilReact == "None React" ~ 0,
                                           rawData_Traumabase$pupilReact == "Unable to assess" | is.na(rawData_Traumabase$pupilReact) ~ -1) 

# Systolic blood pressure
rawData_Traumabase$systolicBloodPressure <- rawData_Traumabase$Pression.Artérielle.Systolique..PAS..à.l.arrivée.du.SMUR 
rawData_Traumabase$systolicBloodPressure <- as.numeric(rawData_Traumabase$systolicBloodPressure)
```



```{r}
summary(rawData_Traumabase[, setdiff(RELEVANT_COVARIATES, "time_to_treatment")])
```



Imputation (already performed and stored in `traumabase_tbi_imputed_mice_for_covariate_selection_paper.RData`)

```{r}
vars.for.imputation <- c("Traitement.anticoagulant",
"Traitement.antiagrégants",
"Glasgow.initial",
"Glasgow.moteur.initial",
"Mannitol...SSH",
"Régression.mydriase.sous.osmothérapie",
"Arrêt.cardio.respiratoire..massage.",
"Fréquence.cardiaque..FC..à.l.arrivée.du.SMUR",
"Cristalloïdes",
"Colloïdes",
"Hémocue.initial",
"Delta.Hémocue",
"Catécholamines",
"SpO2.min",
"Score.de.Glasgow.en.phase.hospitalière",
"Glasgow.moteur",
"Anomalie.pupillaire..Phase.hospitalière.",
"FC.en.phase.hospitalière",
"Doppler.TransCrânien..DTC...Index.de.Pulsatilité..IP..max",
"FiO2",
"Bloc.dans.les.premières.24h....Neurochirurgie..ex....Craniotomie.ou.DVE.",
"Total.Score.IGS",
"Osmothérapie",
"HTIC...25.PIC.simple.sédation.",
"Dérivation.ventriculaire.externe..DVE.",
"Craniectomie.dé.compressive",
"ISS....Head_neck",
"ISS....Face",
"ISS....External",
"Score.ISS",
"ISS....Selection",
"age",
"systolicBloodPressure",
"pupilReact_num",
"gender")

Traumabase.imputed <- missRanger(rawData_Traumabase[, vars.for.imputation], num.trees = 10)
```


```{r}
names(Traumabase.imputed)[names(Traumabase.imputed) == "sexe"] <- "gender"

print(paste0("Raw traumabase data contains following number of observations: ", nrow(Traumabase.imputed)))
```


```{r}
# create a synthetic time to treatment covariates following a beta law
simulated_time_to_treatment <- rbeta(nrow(Traumabase.imputed), 2, 5, ncp = 0)*310
hist(simulated_time_to_treatment)

Traumabase.imputed$time_to_treatment <- simulated_time_to_treatment
```



```{r}
Traumabase.imputed <- Traumabase.imputed[, RELEVANT_COVARIATES]
```


```{r}
Traumabase.imputed$Glasgow.initial <- floor(Traumabase.imputed$Glasgow.initial)
Traumabase.imputed$gender <- as.factor(Traumabase.imputed$gender)
Traumabase.imputed$pupilReact_num <- as.factor(Traumabase.imputed$pupilReact_num)
Traumabase.imputed$Glasgow.initial <- as.factor(Traumabase.imputed$Glasgow.initial)
```


```{r}
summary(Traumabase.imputed)
```


# Binding the two data sets and ensuring overlap

```{r}
Traumabase.imputed$S <- rep(0, nrow(Traumabase.imputed))
CRASH3.imputed$S <- rep(1, nrow(CRASH3.imputed))

total <- rbind(Traumabase.imputed, CRASH3.imputed)
```


## Categorization of the continuous covariates

We need to make categorical some of the covariates that are initially categorical, namely:
- Age
- Systolic Blood Pressure
- Time to treatment

We also lower a bit the granularity of Glasgow score.


```{r}
total$age.categorized <- case_when(total$age < 35 ~ 1,
                                total$age >= 35 &  total$age < 50 ~ 2,
                                total$age >= 50 &  total$age < 60 ~ 3,
                                total$age >= 60 ~ 4)


total$systolicBloodPressure.categorized <- case_when(total$systolicBloodPressure < 120 ~ 1,
                                total$systolicBloodPressure >= 120 &  total$systolicBloodPressure < 180 ~ 2,
                                total$systolicBloodPressure >= 180 ~ 3)


total$time_to_treatment.categorized <- case_when(total$time_to_treatment < 30 ~ 1,
                                total$time_to_treatment >= 30 &  total$time_to_treatment < 60 ~ 2,
                                total$time_to_treatment >= 60 &  total$time_to_treatment < 90 ~ 3,
                                total$time_to_treatment >= 90 &  total$time_to_treatment < 150 ~ 4,
                                total$time_to_treatment >= 150 ~ 5)


total$Glasgow.initial <- as.integer(total$Glasgow.initial) + 2
total$Glasgow.initial <- ifelse(total$Glasgow.initial  == 3, 4, total$Glasgow.initial)
total$Glasgow.initial <- ifelse(total$Glasgow.initial == 15, 14, total$Glasgow.initial)

total$S <- as.factor(total$S)
```


```{r}
summary(total)
```

## Sanity check: Is overlap ensured?

We can count how many observations fall in each categories.

```{r}
count.observations.in.each.strata.trial <- total[total$S == 1,] %>% group_by(time_to_treatment.categorized, Glasgow.initial, systolicBloodPressure.categorized, age.categorized, gender, pupilReact_num) %>%
  summarise(count = n())

count.observations.in.each.strata.target <- total[total$S == 0,] %>% group_by(time_to_treatment.categorized, Glasgow.initial, systolicBloodPressure.categorized, age.categorized, gender, pupilReact_num) %>%
  summarise(count = n())
```

What are the most represented categories in each population?

- In Traumabase, many men, with relative high Glasgow score and good pupils reactivity.
- In trial, same but treated later and a bit more women


```{r}
count.observations.in.each.strata.target[count.observations.in.each.strata.target$count > 0.2*max(count.observations.in.each.strata.target$count), ]

count.observations.in.each.strata.trial[count.observations.in.each.strata.trial$count > 0.2*max(count.observations.in.each.strata.trial$count), ]
```


```{r}
# keep only the data for which we have overlap
total.with.overlap <- left_join(count.observations.in.each.strata.trial, total, by = c('time_to_treatment.categorized', 'Glasgow.initial', 'systolicBloodPressure.categorized', 'age.categorized', 'gender', 'pupilReact_num'), all.y = FALSE, all.x = TRUE)


total.with.overlap <- total.with.overlap[, c('time_to_treatment.categorized', 'Glasgow.initial', 'systolicBloodPressure.categorized', 'age.categorized', 'gender', 'pupilReact_num', 'time_to_treatment', 'age', 'systolicBloodPressure', 'S')]
```

Keeping only observations with overlap leads to a loss of `r nrow(total) - nrow(total.with.overlap)` observations. At the end the final data set contains `r nrow(total.with.overlap)` observations.

```{r}
table(total$S)
table(total.with.overlap$S)
```

Update the count in target population.

```{r}
count.observations.in.each.strata.target <- total.with.overlap[total.with.overlap$S == 0,] %>% group_by(time_to_treatment.categorized, Glasgow.initial, systolicBloodPressure.categorized, age.categorized, gender, pupilReact_num) %>%
  summarise(count = n())
```

## Save data

```{r}
total$pupilReact_num <- as.integer(total$pupilReact_num) - 2

save(total.with.overlap, file = "./data/semi-synthetic-DGP.rds")
save(count.observations.in.each.strata.target, file = "./data/semi-synthetic-oracle-target.rds")
save(count.observations.in.each.strata.trial, file = "./data/semi-synthetic-oracle-trial.rds")
```

## Rename the columns for nice plots

```{r}
total.with.overlap$age.categorized <- case_when(total.with.overlap$age < 35 ~ "< 35",
                                total.with.overlap$age >= 35 &  total.with.overlap$age < 50 ~ "35 - 50",
                                total.with.overlap$age >= 50 &  total.with.overlap$age < 60 ~ "50 - 60",
                                total.with.overlap$age >= 60 ~ "> 60")
total.with.overlap$age.categorized <- as.factor(total.with.overlap$age.categorized)


total.with.overlap$systolicBloodPressure.categorized <- case_when(total.with.overlap$systolicBloodPressure < 120 ~ "Lower range",
                                total.with.overlap$systolicBloodPressure >= 120 &  total.with.overlap$systolicBloodPressure < 180 ~ "Normal range",
                                total.with.overlap$systolicBloodPressure >= 180 ~ "Upper range")

total.with.overlap$systolicBloodPressure.categorized <- as.factor(total.with.overlap$systolicBloodPressure.categorized)


total.with.overlap$time_to_treatment.categorized <- case_when(total.with.overlap$time_to_treatment < 30 ~ "< 30",
                                total.with.overlap$time_to_treatment >= 30 &  total.with.overlap$time_to_treatment < 60 ~ "30 - 60",
                                total.with.overlap$time_to_treatment >= 60 &  total.with.overlap$time_to_treatment < 90 ~ "60 - 90",
                                total.with.overlap$time_to_treatment >= 90 &  total.with.overlap$time_to_treatment < 150 ~ "90 - 150",
                                total.with.overlap$time_to_treatment >= 150 ~ "> 150")

total.with.overlap$time_to_treatment.categorized <- as.factor(total.with.overlap$time_to_treatment.categorized)


total.with.overlap$Glasgow.initial <- as.integer(total.with.overlap$Glasgow.initial) + 2
total.with.overlap$Glasgow.initial <- ifelse(total.with.overlap$Glasgow.initial  == 3, 4, total.with.overlap$Glasgow.initial)
total.with.overlap$Glasgow.initial <- ifelse(total.with.overlap$Glasgow.initial == 15, 14, total.with.overlap$Glasgow.initial)
total.with.overlap$Glasgow.initial <- as.factor(total.with.overlap$Glasgow.initial)


total.with.overlap$pupilReact_num <- as.factor(total.with.overlap$pupilReact_num)
total.with.overlap$gender <- ifelse(total.with.overlap$gender == 1, "Female", "Male")

total.with.overlap$S <- as.factor(total.with.overlap$S)
```



## Looking a bit more to the data - Plots


```{r}
total.with.overlap$age.categorized <- factor(total.with.overlap$age.categorized, order = TRUE, levels = c("< 35", "35 - 50", "50 - 60", "> 60"))

total.with.overlap$time_to_treatment.categorized <- factor(total.with.overlap$time_to_treatment.categorized, order = TRUE, levels =c("< 30", "30 - 60", "60 - 90", "90 - 150",  "> 150"))
```


```{r}
total.with.overlap$Source <- ifelse(total.with.overlap$S == 1, "CRASH-3", "Traumabase")
```


```{r}
n = nrow(total.with.overlap[total.with.overlap$S == 1,])
m = nrow(total.with.overlap) - n

total.with.overlap %>% 
  group_by(time_to_treatment.categorized, Source) %>%  
  summarize(percent = if(Source == "CRASH-3") round(100*n()/n,0)
                          else round(100*n()/m,0)) %>%
  mutate(percent = floor(percent)) %>%
  ggplot(aes(x = time_to_treatment.categorized, fill = Source, group = Source)) +
  geom_col(aes(y = percent), position = "dodge") +
  ylab("%") +
  theme_minimal() +
  xlab("Time between accident and treatment (minutes)") +
  scale_fill_manual(values=c("#E69F00", "#999999"))
  ggsave("./fig/semi-synthetic-ttt.pdf", width = 5, height = 5)

total.with.overlap %>% 
  group_by(systolicBloodPressure.categorized, Source) %>%  
  summarize(percent = if(Source == "CRASH-3") round(100*n()/n,0)
                          else round(100*n()/m,0)) %>%
  mutate(percent = floor(percent)) %>%
  ggplot(aes(x = systolicBloodPressure.categorized, fill = Source, group = Source)) +
  geom_col(aes(y = percent), position = "dodge") +
  ylab("%") +
  theme_minimal() +
  xlab("Systolic blood pressure") +
  scale_fill_manual(values=c("#E69F00", "#999999"))
ggsave("./fig/semi-synthetic-blood.pdf", width = 5, height = 5)

total.with.overlap %>% 
  group_by(age.categorized, Source) %>%  
  summarize(percent = if(Source == "CRASH-3") round(100*n()/n,0)
                          else round(100*n()/m,0)) %>%
  mutate(percent = floor(percent)) %>%
  ggplot(aes(x = age.categorized, fill = Source, group = Source)) +
  geom_col(aes(y = percent), position = "dodge") +
  ylab("%") +
   theme_minimal() +
  xlab("Age (years)") +
  scale_fill_manual(values=c("#E69F00", "#999999"))
ggsave("./fig/semi-synthetic-age.pdf", width = 5, height = 5)

total.with.overlap %>% 
  group_by(gender, Source) %>%  
  summarize(percent = if(Source == "CRASH-3") round(100*n()/n,0)
                          else round(100*n()/m,0)) %>%
  mutate(percent = floor(percent)) %>%
  ggplot(aes(x = gender, fill = Source, group = Source)) +
  geom_col(aes(y = percent), position = "dodge") +
  ylab("%") +
   theme_minimal() +
  #xlab("Gender") +
  scale_fill_manual(values=c("#E69F00", "#999999"))
ggsave("./fig/semi-synthetic-gender.pdf", width = 5, height = 5)

total.with.overlap %>% 
  group_by(Glasgow.initial, Source) %>%  
  summarize(percent = if(Source == "CRASH-3") round(100*n()/n,0)
                          else round(100*n()/m,0)) %>%
  mutate(percent = floor(percent)) %>%
  ggplot(aes(x = Glasgow.initial, fill = Source, group = Source)) +
  geom_col(aes(y = percent), position = "dodge") +
  ylab("%") +
  theme_minimal() +
  xlab("Glasgow score") +
  scale_fill_manual(values=c("#E69F00", "#999999"))
ggsave("./fig/semi-synthetic-glasgow.pdf", width = 5, height = 5)


total.with.overlap %>% 
  group_by(pupilReact_num, Source) %>%  
  summarize(percent = if(Source == "CRASH-3") round(100*n()/n,0)
                          else round(100*n()/m,0)) %>%
  mutate(percent = floor(percent)) %>%
  ggplot(aes(x = pupilReact_num, fill = Source, group = Source)) +
  geom_col(aes(y = percent), position = "dodge") +
  ylab("%") +
   theme_minimal() +
  xlab("Pupil reactivity") +
  scale_fill_manual(values=c("#E69F00", "#999999"))
ggsave("./fig/semi-synthetic-pupil.pdf", width = 5, height = 5)

```

```{r}
ggplot(total.with.overlap, aes(x = time_to_treatment, fill = Source, group = Source)) +
  geom_density(alpha = 0.5) +
  theme_classic()
```


